---
title: "Debugging Parallel stringdist"
author: "Peter von Rohr"
date: '2018-05-17'
output:
  html_document:
    df_print: paged
---


## Disclaimer
The R-script `GPsearch_stringdist-hamming-parallel-singleAnimal.R` should run an all-against-all comparison of SNP-strings. This document shows some steps in the process of debugging some problems of the script.


## Current Status
* Running the script on a small dataset of only 1000 animals to be compared runs fine, because only one comparison iteration has to be peformed.
* Comparing all-against-all gives the following errror

```
Error in { : task 10 failed - "subscript out of bounds"
Calls: %dopar% -> <Anonymous>
Execution halted
```


## Debugging
The error points to a problem with computing index bounds. Hence, we start by computing some of these index boundaries.

```{r}
# number of records read
NORECS <- 97970
STRATA <- "ALLANIMALS"
```

Constants defining limitation value

```{r}
const_limit <- 10^8
foreach_chunk_size <- 10
```

```{r}
# if(STRATA=="ALLANIMALS"){if((length(gta)^2) <= const_limit){limitation<-length(gta)^2}else{limitation<-const_limit}}
 if(STRATA=="ALLANIMALS"){if((NORECS^2) <= const_limit){limitation<-NORECS^2}else{limitation<-const_limit}}
```

As a result we changed limitation to 

```{r}
cat("limitation: ", limitation)
```

The variable `total_number` stands for total number of comparison that must be done in  our all-against-all

```{r}
nr_elemente <- as.numeric(NORECS) 
if(STRATA == "ALLANIMALS"){
  total_number <- nr_elemente^2
}
if(STRATA == "SINGLEANIMAL"){
  total_number <- nr_elemente*length(selani)
}
cat("Total number: ", total_number, "\n")
```

Due to limitation in RAM or in R, we have to split the total number of comparisons into chunks of possible comparisons. Each chunk can have `nr_possible_comparisons` of comparisons.

```{r}
nr_possible_comparisons <- limitation
```


```{r}
nr_loops <- floor(total_number / nr_possible_comparisons)
cat("NR LOOPS: ", nr_loops, "\n")
```

There might be some comparisons left after all the loops. The number of remaining comparisons is

```{r}
total_number - nr_loops * nr_possible_comparisons
```


```{r}
nr_chunks_per_iter <-  floor(nr_possible_comparisons / nr_elemente)
cat("nr chunks: ", nr_chunks_per_iter, "\n")
```

The remaining number of chunks

```{r}
nr_possible_comparisons - nr_chunks_per_iter * nr_elemente
```


### Unrolling outer loop
The comparisons that are done are defined by a matrix with  two rows. This is constructed from two vectors `base_vec` and `first_vec`. 

```{r}
chunk_idx <- 1
base_vec <- rep(1:nr_elemente,nr_chunks_per_iter)
head(base_vec, n=20)
tail(base_vec, n=20)

  if (STRATA=="ALLANIMALS"){
    first_vec <- rep(((chunk_idx-1)*nr_chunks_per_iter + 1):(chunk_idx * nr_chunks_per_iter), nr_elemente)
  }
  if (STRATA=="SINGLEANIMAL"){
    first_vec <- rep(selani[1]:selani[2],nr_elemente)
  }
head(first_vec[order(first_vec)], n=20)
tail(first_vec[order(first_vec)], n=20)

  cat("first_vec created ", " ", format(Sys.time(), "%a %b %d %X %Y"), "\n")
```

Composing the matrix that must have the following number of columns

```{r}
nr_chunks_per_iter * nr_elemente
```


```{r}
m_result <- matrix(c(first_vec[order(first_vec)], base_vec), nrow = 2, byrow = TRUE)
m_result[,1:20]
dim(m_result)
m_result[,(ncol(m_result)-20):ncol(m_result)]
```

Inner parallelisation using `foreach`. First starting positions are defined in `ii`

```{r}
ii = seq(1,nr_possible_comparisons,nr_possible_comparisons/foreach_chunk_size)
ii = ii[1:(length(ii)-1)]
length(ii)
cat("starting positions: ")
print(as.integer(ii))
```


```{r}
as.integer(ii+((nr_possible_comparisons/foreach_chunk_size) - 1))
```

New version

```{r}
startingPoints <- seq(1,nr_possible_comparisons,nr_possible_comparisons/foreach_chunk_size)
ii = startingPoints[1:(length(startingPoints)-1)]
cat("starting positions: ")
print(as.integer(ii))

```

```{r}
as.integer(ii+((nr_possible_comparisons/foreach_chunk_size) - 1))
```


The remaining comparisons

```{r}
as.integer(startingPoints[length(startingPoints)])
```

The last is

```{r}
ncol(m_result)
```



```{r, echo=FALSE}
cat(rmddochelper::get_latest_change(), "\n")
```

