---
title: "Experiments In Parallelisation In R"
subtitle: "Using stringdist as an example"
author: "Peter von Rohr"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```


# Disclaimer
This document describes a first set of experiments with parallelisation in R. We are using pair-wise comparisons between strings using the R-package `stringdist`^[GitHub https://github.com/markvanderloo/stringdist]. There is a paper on stringdist^[R-Journal https://journal.r-project.org/archive/2014-1/loo.pdf]


# Introduction
There are two motivations behind the experiments shown in here

1. string comparisons is a hard and an important problem
2. when it comes to increasing the preformance above a certain level, there is no way around parallelisation.


# Background
## String comparisons
The background and the importance of string comparison is described in [@vanderLoo2014]. Furthermore, the cited paper also describes a list of R-packages in the context of string comparison, are given. 

## Parallelisation
The technology of parallelisation used in `stringdist` is multithreading based on openmp^[On line course taught by Tom Mattson from Intel, https://www.youtube.com/playlist?list=PLLX-Q6B8xqZ8n8bwjGdzBJ25X2utwnoEG]


# Experiments
This section describes concrete applications of the `stringdist` functionality to different types of data. The focus of the experiments are on different ways of parallelisation of the computations needed for comparing strings using stringdist.

An important question is whether the built-in parallelisation functionality works, compared to other parallelisation frameworks. What is already clear is that the built-in parallelisation of stringdist will not work on MacOS (and we are not sure about the situation in Windows), because the underlying parallelisation framework - openmp - used in stringdist is not available in `clang` which is the default compiler on MacOS, nowadays.


```{r bib_section, echo=FALSE}
s_bib_file <- "skeleton.bib"
vec_bref <- c(bibentry(
  bibtype = "Article",
  title = "The stringdist Package for Approximate String Matching",
  author = as.person("M. P.J. van der Loo"),
  journal = "R Journal",
  url = "https://journal.r-project.org/archive/2014-1/loo.pdf",
  volumne = "6/1",
  year = "2014",
  key = "vanderLoo2014"
))

rmddochelper::write_bib(pvec_bref = vec_bref)
```


